---
title: "Pipeline example"
format: html
engine: julia
execute:
    echo: true
    eval: false
julia: 
    exeflags: ["--project=../CoEvolution"]
---

# Introduction

This file shows how to utilize the pipeline built here to generate an ERC network for some species of interest. In this case, we will be focusing on *D. melanogaster* and utilizing all annotated Dipteran genomes available on ncbi, alongside *C. felis* as an outgroup. We assume that you are using some form of Unix, but the pipeline has been tested on Ubuntu running in WSL2. It *should* work in both Windows and on MacOS, assuming all necessary software is installed and visible to your CLI.

# Initializing the software, installing all dependencies.

Before we run any analyses, we will require several tools. 1) Julia v 1.11+

Once Julia is installed, run the below script:

```{julia}
using Pkg
Pkg.add("DrWatson")
Pkg.activate("CoEvolution")
Pkg.instantiate()
```

This might run for a while, as it will compile any necessary packages for later scripts. While that is running, install the following on your system.

2)  [Mafft](https://github.com/rcedgar/muscle/releases)
3)  [Diamond](https://github.com/bbuchfink/diamond/wiki)
4)  [iqTree2](http://www.iqtree.org/#download)

If you would like to use this pipeline to download and prep data, make sure you have installed:

5)  [datasets-cli](https://www.ncbi.nlm.nih.gov/datasets/docs/v2/command-line-tools/download-and-install/)
6)  [ncbi e-utils](https://www.ncbi.nlm.nih.gov/books/NBK179288/)
7)  seq-kit

Hopefully Julia is up and running at this point, so we can begin on the main steps of the analysis.

# Defining major parameters

The below codeblock is where you can define the major parameters for the pipeline (taxon, focal species, outgroup, directory). The nature of this pipeline means it might be a good idea to just have these in the `Project.toml` file, but I know people will be more afraid of editing a file they can't see versus running code-blocks in a somewhat familiar interface.

```{julia}
#| label: initialization
#| echo: true

using DrWatson
@quickactivate "CoEvolution"

#Set these values for your own analysis. Taxon can be given as plain-text, but use ncbi taxon ids when possible.
taxon = "Diptera"
data_dir = datadir("dipterans/") #The default place for all of the files, but give the absolute path of wherever data is meant to go otherwise.
#Specify the focal species for the analysis.
focal= "D_melanogaster"
#Specify an outgroup - necessary for ASTRAL-IV to adequately make a species tree.
outgroup_ncbi_id = "Ctenocephalides felis"
outgroup_name = "C_felis"
include(srcdir("Stats_utilities.jl"))
include(srcdir("Phylo_utilities.jl"))
```

# Prep data

You can prepare the data in one of two ways. First, you can define the taxon of interest in the code block below and download data using our internal script. This will poll the NCBI `datasets` database to find any annotated, chromosome level assemblies within the taxon specified, download their protein sequences and gff files and place them in the data directory in the `seqs` folder.

```{julia}
#| label: data download
#| 
#SKIP THIS IF YOU ARE USING YOUR OWN DATA!
include(srcdir("Finding_orthologs.jl"))
isdir(data_dir) || mkdir(data_dir)
download_and_prep_sequences(taxon)
download_and_prep_sequences(outgroup_ncbi_id;level="scaffold")
```

# Finding orthologs for focal species.

We first identify orthologous sequences for all genes within the focal species. This is done in the following code block, most functions are defined in the `src\Finding_orthologs.jl` script. Fill out parameters in the blanks below to point to the correct files.

```{julia}
#| label: import ERC and Ortholog functions
include(srcdir("Finding_orthologs.jl"))
include(srcdir("ERC_functions.jl"))
```

```{julia}
#| label: find orthologs
#Now run the below, and go off to grab some tea, or take a break for the day.
species = replace.(filter(endswith(".faa"),readdir(data_dir*"seqs/")),r".faa"=>"")
all_orthologs = find_orths(focal,species,data_dir)
#If the above gets interrupted - delete any .tsv files of size 0 in your `data_dir\blast` directory, and run it again. It will only run analyses for missing pairs of species.
using Arrow

Arrow.write(data_dir*focal*"_orthologs.arrow",all_orthologs)

#And if things break later, you can always load in the all_orthologs dataframe with `all_orthologs=DataFrame(Arrow.Table(data_dir*focal*"_orthologs.arrow))"`

#The next step takes a while, since parallelization is doable but not recommended (mafft can take quite a bit of memory, so expect slow downs)

build_orth_files(all_orthologs,data_dir)

#Finally, we generate a tree for each gene. Earlier versions we constrained the tree topology, which was faster, but does not allow us to do certain secondary analyses that we will want to, and is probably the wrong way to approach this.

using FLoops
using ProgressMeter

isdir(data_dir*"trees/") || mkdir(data_dir*"trees/")

aligned_genes = filter(endswith(".aligned.fasta"),readdir(data_dir*"aligned/"))

p=Progress(length(aligned_genes),desc="Running IQTree2")
@floop ThreadedEx() for seq in aligned_genes
    try 
        run_iqtree(seq)
    catch    
    end
    next!(p)
end

finished_trees = filter(endswith(".treefile"),readdir(data_dir*"trees/",join=true))
run_astral(finished_trees)

```

## Calculating ERC

If you are using your own trees, start here!

You will need two specify the species tree and the list of trees to compare.

```{julia}
#| label: ERC calculations
#Change if using some other species tree.
species_tree = read_tree(data_dir*"trees/astral_consensus_rerooted.newick")
tree_files = filter(endswith(".treefile"),readdir(data_dir*"trees/",join=true))
#This takes a while, but saves a lot of reading time in the ERC loop. It does increase memory requirements even further, consider running `runERC_files(trefiles,species_tree)` if you are being memory limited. But, you are avoiding sum(1:length(tree_files)) file read operations, which is also probably a benefit.
trees = [read_tree(t) for t in tree_files]

ERC = runERC_collection(trees,species_tree;cutoff=5,min_shared=4)
#Drop cases with missing data 
ERC = ERC[findall(completecases(ERC)),:]
genes = replace.(filter(endswith(".treefile"),readdir(data_dir*"trees/",join=false)),r".treefile"=>s"")
ERC.i = genes[Int.(ERC.i)]
ERC.j = genes[Int.(ERC.j)]
ERC.gid = [join(sort([x.i,x.j]),"_") for x in eachrow(ERC)]
using Arrow
Arrow.write(data_dir*focal*"_ERC.arrow",ERC)
ERC = DataFrame(Arrow.Table(data_dir*focal*"_ERC.arrow"))
```

## Gene Table Annotation and comparisons

We next build a table that includes all of the gene information. This is based off of the .gff file for your reference of interest.

The chr_dict is just a quick conversion from NCBIs chr ids to something human readable - you'll have to make your own for your system. Note, we drop anything that's on an unmapped scaffold in the process, in our case 6 genes. If you have hundreds of chromosomes/scaffolds, this step can be skipped, or you can use `Julia`s ability to make Dicts iteratively.

```{julia}
#| label: gene_table import
include(srcdir("Annotation_functions.jl"))

genes = replace.(filter(endswith(".treefile"),readdir(data_dir*"trees/",join=false)),r".treefile"=>s"")

gene_table = DataFrame(gene=genes,
    stable_id=missings(String,length(genes)),
    name=missings(String,length(genes)),
    flybase=missings(String,length(genes)),
    chr=missings(String,length(genes)),
    pos_1=missings(Int64,length(genes)),
    pos_2=missings(Int64,length(genes)),
    description=missings(String,length(genes)),
    num_rbh=missings(Int64,length(genes)))

#Can't query all 12k at once, split into chunks of 100
part_size=200
parts = collect(Iterators.partition(genes,part_size))
for gs in 1:length(parts)
    gene_names = copy(parts[gs])
    subtable = find_name(gene_names,"dmelanogaster_gene_ensembl")
    rows = indexin(subtable.gene,genes)
    gene_table[rows,2:8] = subtable[:,2:8]
    gene_table.num_rbh[(1:length(gene_names)).+part_size*(gs-1)] = num_rbh.(gene_names)
end

#For genes with no Ensembl annotation, we look in the gff file to at least get location data

using CSV

gff_table = CSV.read(data_dir*focal*"_genes.gff",DataFrame,header=false,skipto=9)
gff_table = gff_table[completecases(gff_table),:]

chr_dict = Dict("NC_004354.4"=>"X","NT_033779.5"=>"2L","NT_033778.4"=>"2R","NT_037436.4"=>"3L","NT_033777.3"=>"3R","NC_004353.4"=>"4","NC_024512.1"=>"Y","NC_024511.2"=>"MT")

gff_table.Column1 = [get(chr_dict,c,"NA") for c in gff_table.Column1]
to_do_genes = findall(ismissing.(gene_table.chr))
for g in to_do_genes
    subtable = gff_table[findall(occursin.(genes[g],gff_table.Column9)),:]
    if size(subtable)[1] > 0
        gene_table.chr[g] = subtable.Column1[1]
        gene_table.pos_1[g] = minimum(subtable.Column4)
        gene_table.pos_2[g] = maximum(subtable.Column5)
    end
end


CSV.write(data_dir*focal*"_gene_table.csv",gene_table)
```

## ERC network analysis

### Finding significant edges

```{julia}
include(srcdir("ERC_simulations.jl"))
species_tree = read_tree(data_dir*"trees/astral_consensus_rerooted.newick")


#Positive null - trees that _have_ evolved along the exact same rates, so should show correlations with minor differences.
trees_null = simulate_tree_set(species_tree,5,1000,4,73,focal)
ERC_sims = runERC_collection(trees_null,species_tree;cutoff=5)
ERC_sims = ERC_sims[completecases(ERC_sims),:]
disallowmissing!(ERC_sims)
sims_cdf = ecdf(ERC_sims.fERC)
ERC.sims = sims_cdf.(ERC.fERC)

#Negative null - correlations between i.i.d. branch lengths drawn from exponential proportional to the branch lengths.
ERC_null = calculate_ERC_exp_null(species_tree,1000)
disallowmissing!(ERC_null)
ERC_null.fERC = fisher_trans.(ERC_null.r,length(get_branch_lengths(species_tree)))
null_cdf = ecdf(ERC_null.fERC)
ERC.null = null_cdf.(ERC.fERC)

#Take only edges outside of expectation from positive sim
gt_sims = findall(ERC.sims .>0.95)
#Now, edges that are Bonferonni significant _and_ beyond dynamic cut-off are retained further.
sig_ERC = ERC[intersect(findall(ERC.sims .> 0.95),findall(ERC.pval .< 0.05/length(gt_sims))),:]

CSV.write(data_dir*focal*"_sig_ints.csv",sig_ERC)
```

### Broad analysis of scores

We start by looking at the broad set of ERC scores. How do they compare to some potential theoretical nulls, and do they map to any interesting gene features?

```{julia}
gene_table = CSV.read(data_dir*focal*"_gene_table.csv",DataFrame)
using Arrow
ERC = DataFrame(Arrow.Table(data_dir*focal*"_ERC.arrow"))

id_i = indexin(ERC.i,gene_table.gene)
id_j = indexin(ERC.j,gene_table.gene)

ERC.chr_i = gene_table.chr[id_i]
ERC.chr_j = gene_table.chr[id_j]

ERC.pos_i_1 = gene_table.pos_1[id_i]
ERC.pos_i_2 = gene_table.pos_2[id_i]

ERC.pos_j_1 = gene_table.pos_1[id_j]
ERC.pos_j_2 = gene_table.pos_2[id_j]

ERC.dist = missings(Int64,length(ERC.i))
intra = findall(ERC.chr_i .== ERC.chr_j)
ERC.same_chr .= "different"
ERC.same_chr[intra] .= "same"
ERC.dist[intra] = [range_dist([x.pos_i_1,x.pos_i_2],[x.pos_j_1,x.pos_j_2]) for x in eachrow(ERC[intra,:])]

Arrow.write(data_dir*focal*"_ERC_annotated.arrow",ERC)
```

Next, some summaries for the gene_table:

```{julia}
#Annotate the gene table.sig_
function indexInERC(gene); findall(ERC.i .== gene .|| ERC.j .== gene);end
idx = map(indexInERC,gene_table.gene)
gene_table.sig_ints = [length(union(findall(sig_ERC.i.==x),findall(sig_ERC.j.==x))) for x in gene_table.gene]
gene_table.mean_r = [mean(ERC.r[idx[i]]) for i in 1:length(idx)]
gene_table.mean_fERC = [mean(ERC.fERC[idx[i]]) for i in 1:length(idx)]
gene_table.mean_shared = [mean(ERC.n_edges[idx[i]]) for i in 1:length(idx)]
CSV.write(data_dir*focal*"_annotated_gene_table.csv",gene_table)
```

Now we generate some basic summary plots for the data:

```{julia}
#| label: fig1
ERC = DataFrame(Arrow.Table(data_dir*focal*"_ERC_annotated.arrow"))

using CairoMakie, LaTeXStrings
include(srcdir("Plot_utilities.jl"))
isdir(data_dir*"plots") || mkdir(data_dir*"plots")
plot_prefix=data_dir*"plots/"*focal*"_"

#Panel A: Histograms of fERC for three distributions
fig1 = Figure(size=(800,800))
ax1 = Axis(fig1[1,1],
    xlabel="fERC",
    ylabel="Frequency",
    xgridvisible=false,
    ygridvisible=false)

hist!(ax1,ERC_sims.fERC,bins=100,
       label="Positive Simulations",normalization=:pdf,
       strokecolor=:purple,color=(:purple,0.5))

hist!(ax1,ERC_null.fERC,bins=100,
       label="Null Expectation",normalization=:pdf,
       strokecolor=:orange,color=(:orange,0.5))

hist!(ax1,ERC.fERC,bins=200,
    label="Real Data",normalization=:pdf,
       strokecolor=:blue,color=(:blue,0.5))
vlines!(ax1,quantile(ERC_null.fERC,[0.95]),color=:orange)
vlines!(ax1,quantile(ERC_sims.fERC,[0.95]),color=:purple)

xlims!(ax1,(-5,15))


axislegend(ax1,merge=true,unique=true)

ax2, hm = hexbin(fig1[1,2],ERC.n_edges,ERC.fERC,
    colormap=[Makie.to_color(:transparent);Makie.to_colormap(:viridis)],
    axis= (; xgridvisible=false,ygridvisible=false,
             ylabel="fERC",xlabel="Shared Edges"),
    colorscale=Makie.pseudolog10,
    bins=(50,50),
    threshold=2)


#loess_1_2 = loess(collect(ERC.n_edges),collect(ERC.fERC),span=0.1) 
#xs = range(5,120,length=200)
#ys = predict(loess_1_2,xs)
#lines!(ax2,xs,ys,linewidth=8,color=:white)
#lines!(ax2,xs,ys,linewidth=4,color=:black)


Colorbar(fig1[1,2][1,2],hm,ticks=[10,100,1000,10000,100000,1000000])

ax3 = Axis(fig1[2,1],yticks=([0.25,0.5],["",""]),xlabel="fERC",ylabel="Pair on _ chromosomes")

sferc = [ERC.fERC[findall(ERC.same_chr.=="same")],ERC.fERC[findall(ERC.same_chr.=="different")]]
for i in 1:2
    d = density!(ax3,sferc[i],offset=i./4,color=:x,colormap=:viridis,colorrange=[-5,15],label=["Same","Different"][i],
    strokecolor=:black,strokewidth=1)
    translate!(d,0,0,-0.1i)
end
xlims!(ax3,(-5,15))
annotation!(ax3,[(1,1/4),(1,1/2)],text=["Same","Different"],
    textcolor=:white,
    fontsize=20)

sub_ERC = findall((!ismissing).(ERC.dist))

ax4,hm = hexbin(fig1[2,2],ERC.dist[sub_ERC],ERC.fERC[sub_ERC],
           colorscale=Makie.pseudolog10,
           bins=(50,50),
           axis=(; xlabel="Pair Distance (bp)",
                  ylabel="fERC Score",
                  xscale=Makie.pseudolog10,
                  xticks=([0,10,1000,100000,10000000],[L"0",L"10",L"10^3",L"10^5",L"10^7"]),
                  xgridvisible=false,ygridvisible=false))
Colorbar(fig1[2,2][1,2],hm,ticks=[10,100,1000,10000,100000,1000000])

lm_1 = lm(@formula(fERC~dist),ERC[sub_ERC,:])
xs = range(extrema(ERC.dist[sub_ERC])...,length=200)
ys = predict(lm_1,DataFrame(dist=collect(xs)))
lines!(ax4,xs,ys,linewidth=8,color=:white)
lines!(ax4,xs,ys,linewidth=4,color=:black)


for (label,layout) in zip(["A","B","C","D"],[fig1[1,1],fig1[1,2],fig1[2,1],fig1[2,2]])
    Label(layout[1,1,TopLeft()],label,
        fontsize=26,
        font=:bold,
        padding=(0,5,5,0),
        halign=:right)
end

save(plot_prefix*"fig1.pdf",fig1)
```

### Chromosome plot

Julia has no good backend for this (well, neither does R), but I want to plot each pair as a point in a heatmap, with the center of each gene. The issue is converting the location to something meaningful across chromosomes.

Need to have the order of genes consistent between x and y, while accounting for different chromosomes. We'll make dummy X and Y coords here, with the order: 2L,2R,3L,3R,4,X,Y,MT

```{julia}
#Pre-compute a bunch of locations/bins
pos_i = (ERC.pos_i_1 .+ ERC.pos_i_2)./2
pos_j = (ERC.pos_j_1 .+ ERC.pos_j_2)./2


chr_order = ["2L","2R","3L","3R","4","X","Y","MT"]
chr_lengths = [23312347,25254477,28012551,32067996,1271129,23107330,3666928]
chr_lines = [sum(chr_lengths[1:x]) for x in 1:7]
chr_label_pos = vcat(chr_lengths[1]/2,[chr_lengths[i]/2+chr_lines[i-1] for i in 2:7])
chr_ticks = (chr_label_pos,chr_order[1:7])

pos_i = chr_order_pos(pos_i,ERC.chr_i,chr_order,chr_lengths)
pos_j = chr_order_pos(pos_j,ERC.chr_j,chr_order,chr_lengths)


#Only plot triangle
pos_map1= [minimum([pos_i[x],pos_j[x]]) for x in 1:length(pos_i)]
pos_map2= [maximum([pos_i[x],pos_j[x]]) for x in 1:length(pos_i)]

#We don't actually want to plot 80 million points, instead we summarize into bins

wnd_size=100000

binned_ERC = binned_mean(DataFrame(p1=pos_map1,p2=pos_map2,r=ERC.r),wnd_size,wnd_size)

#Now locations of significant interactions
pos_i = (sig_ERC.pos_i_1 .+ sig_ERC.pos_i_2)./2
pos_j = (sig_ERC.pos_j_1 .+ sig_ERC.pos_j_2)./2

pos_i = chr_order_pos(pos_i,sig_ERC.chr_i,chr_order,chr_lengths)
pos_j = chr_order_pos(pos_j,sig_ERC.chr_j,chr_order,chr_lengths)

#Only plot lower triangle
pos_map2= [minimum([pos_i[x],pos_j[x]]) for x in 1:length(pos_i)]
pos_map1= [maximum([pos_i[x],pos_j[x]]) for x in 1:length(pos_i)]


gene_pos = chr_order_pos((gene_table.pos_1.+gene_table.pos_2)./2,gene_table.chr,chr_order,chr_lengths)

# And the actual plotting

fig2 = Figure()

#Mean gene fERC plot
ax1 = Axis(fig2[1,1],xlabel="fERC",ylabel="Position",
    yticks=chr_ticks,
    yticksize=0,
    yminorticks=chr_lines,
    yminorticksvisible=true,
    yminorgridvisible=true,
    yminorgridcolor=RGBAf(0,0,0,0.25),
    ygridcolor=RGBAf(0,0,0,0),
    limits =(nothing,nothing,0,maximum(chr_lines)))
scatter!(ax1,gene_table.mean_fERC,gene_pos,color=gene_table.sig_ints,markersize=3)
loess1 = loess(gene_pos,gene_table.mean_fERC,span=0.1)
xs=range(extrema(gene_pos)...;step=100000)
ys = predict(loess1,xs)
lines!(ax1,ys,xs,linewidth=8,color=:white)
lines!(ax1,ys,xs,linewidth=4,color=:black)

#Genome-wide interaction plot
ax2 = Axis(fig2[1,2],xlabel="",ylabel="", 
    yticklabelsvisible=false,
    xticklabelsvisible=false,
    xminorticks=chr_lines,
    yminorticks=chr_lines,
    xticksize=0,
    yticksize=0,
    xminorticksvisible=true,
    yminorticksvisible=true,
    yminorgridvisible=true,
    yminorgridcolor=RGBAf(0,0,0,0.25),
    ygridcolor=RGBAf(0,0,0,0),
    xminorgridvisible=true,
    xminorgridcolor=RGBAf(0,0,0,0.25),
    xgridcolor=RGBAf(0,0,0,0)
)
heatmap!(ax2,binned_ERC.x_bin,binned_ERC.y_bin,binned_ERC.r_mean)
scatter!(ax2,pos_map1,pos_map2,color=:red,markersize=1)



#Significant interaction plot
ax3 = Axis(fig2[2,2],xlabel="Position",ylabel="Degree",
    xticks=chr_ticks,
    xticksize=0,
    xminorticks=chr_lines,
    xminorticksvisible=true,
    xminorgridvisible=true,
    xminorgridcolor=RGBAf(0,0,0,0.25),
    xgridcolor=RGBAf(0,0,0,0),
    yticks=[1,400,800,1200],
    limits =(0,maximum(chr_lines),nothing,nothing))
scatter!(ax3,gene_pos,gene_table.sig_ints,color=gene_table.sig_ints,markersize=2)

#Resize for emphasis on interaction plot
rowsize!(fig2.layout,2,Auto(0.2))
colsize!(fig2.layout,1,Auto(0.2))
colgap!(fig2.layout,-15)
rowgap!(fig2.layout,-15)

save(plot_prefix*"fig2.pdf",fig2)

```

## Community detection

We next detect communities of genes based on co-evolutionary interactions. We use the Louvain clustering approach, which tries to maximize modularity within each community (e.g. how dense interactions are within each community versus out of it). We use the Leiden algorithm approach implemented in R. We can try to get `R` and `Julia` to interoperate using `RCall`, but we've found in testing this is unreliable for some users. Please modify the location of your files in the below code.

```{r}
require(igraph)
require(readr)

sig_ERC = read_csv("data/dipterans/D_melanogaster_sig_ints.csv")
edges = c(apply(sig_ERC,1,function(x) c(x[1],x[2])))
graph = make_graph(edges,directed=FALSE)

weights = sig_ERC$fERC
weights = weights - min(weights)
weights = weights / max(weights)
E(graph)$weight = weights

#Identify only larges connected component
comps = components(graph)
drop = names(comps$membership)[which(comps$membership != 1)]
subgraph = delete_vertices(graph,drop)

comms = cluster_leiden(subgraph,objective_function="modularity",resolution=1,weights = E(subgraph)$weight,n_iterations=10000)
                
comms$membership = match(1:max(comms$membership),order(table(comms$membership),decreasing=TRUE))[comms$membership]

write_tsv(data.frame(name=comms$names,community=comms$membership),file="data/dipterans/D_melanogaster_communities.tsv")
```

## GO analysis

The resulting communities are examined for GO category enrichment. The easiest approach here is to use `gprofiler2`, which has a nice implementation in R, so once again we'll step out of Julia for a bit, using `RCall` to make the calls from inside julia. You will need `R` and `gprofiler2` installed for this to work. Alternatively, you can run the analysis in `R`, but it will require some data transfer

```{r}
require(gprofiler2)
require(ggplot2)
require(GOSemSim)
require(org.Dm.eg.db)
require(clusterProfiler)
require(emmapplot)


gene_table = read_csv("data/dipterans/D_melanogaster_annotated_gene_table.csv")
#Five major clusters
big_comms = which(table(comms$membership)>50)
cs = lapply(1:length(big_comms),function(x){
    genes = comms$names[which(comms$membership == x)]
    idx = match(genes,gene_table$gene)
    stable_ids = gene_table$stable_id[idx]
    stable_ids = stable_ids[!is.na(stable_ids)]
    stable_egs = mapIds(org.Dm.eg.db,keys=stable_ids,column="ENTREZID",keytype="FLYBASE",multiVals="first")
    return(stable_egs)
})
names(cs) = sapply(1:length(big_comms),function(x) paste("comm_",x,sep=""))

genes=gsub(".treefile","",list.files("data/dipterans/trees",pattern=".treefile"))
stable_genes = sapply(genes,function(x) gene_table$stable_id[which(gene_table$gene == x)][1])
stable_EGs = mapIds(org.Dm.eg.db,keys=stable_genes,column="ENTREZID",keytype="FLYBASE",multiVals="first")           
                      
gostres = gost(cs,organism="dmelanogaster",custom_bg=genes,ordered_query=FALSE,highlight = TRUE)
write_csv(gostres$result,file="data/dipterans/D_melanogaster_fullGOProfiler.csv")
#After inspection of results, we chose some top terms to highlight for export tables, each table saved seperately

highlighted = gostres$result[which(gostres$result$highlighted),]
write_csv(highlighted,file="data/dipterans/D_melanogaster_highlighted.csv")

gene_table$community = sapply(gene_table$gene,function(x) comms$membership[which(comms$name==x)][1])
                                                     
gene_table$comm_degree = NA
#Annotate by degree w/in community
for( x in 1:length(big_comms)) {
    idx = which(gene_table$community == big_comms[x])
    subERC = subset(sig_ERC,i %in% gene_table$gene[idx] & j %in% gene_table$gene[idx])
    for(i in 1:length(gene_table$gene[idx])){
        gene_table$comm_degree[idx][i] = length(union(which(subERC$i == gene_table$gene[idx][i]),which(subERC$j == gene_table$gene[idx][i])))
    }
}

write_csv(gene_table,"data/dipterans/D_melanogaster_annotated_gene_table.csv")

emmap_cluster = function(gene_list,cats=30) {
    ego = enrichGO(gene=gene_list,universe=as.character(stable_EGs),OrgDb=org.Dm.eg.db,ont="BP",pAdjustMethod="BH",pvalueCutoff=0.01,readable=TRUE)
    edo = pairwise_termsim(ego)
    return(emapplot(edo,showCategory=cats))
}


gene_predict = function(gene) {
    subERC = subset(sig_ERC,i==gene | j == gene)
    partners = setdiff(union(subERC$i,subERC$j),gene)
    partner_ids = gene_table$stable_id[match(partners,gene_table$gene)]
    return(enrichGO(partner_ids,keyType="FLYBASE",
                   OrgDb = org.Dm.eg.db,
                   universe = gene_table$stable_id,
                   ont="ALL",
                   pvalueCutoff=0.01,
                   pAdjustMethod="BH"))
}

                              
#Figure SX
obs = table(sig_ERC$chr_combo)
expo = sapply(names(obs),function(x){
    chrs=strsplit(x,split="-");
    scale=1;
    if(chrs[[1]][[1]]==chrs[[1]][[2]]){
        choose(length(which(gene_table$chr==chrs[[1]][[1]])),2)   
    } else {
        length(which(gene_table$chr==chrs[[1]][1]))*length(which(gene_table$chr==chrs[[1]][2]))    
    }
})
perc = length(sig_ERC$r)/(choose(length(gene_table$gene),2))
expo = expo * perc
rat = obs/expo
dev_table = tibble(obs=as.numeric(obs),exp=floor(as.numeric(expo)),ratio=as.numeric(rat))
chrs = strsplit(names(obs),split="-")
dev_table$chr_i = unlist(lapply(chrs,function(x) x[1]))
dev_table$chr_j = unlist(lapply(chrs,function(x) x[2]))

require(ggplot2)
require(cowplot)
theme_set(theme_cowplot())

chr_plot = ggplot(dev_table,aes(x=chr_i,y=chr_j))+geom_tile(aes(fill=ratio))+
            scale_fill_continuous(type="viridis",trans="log2",breaks=c(0.25,0.5,1,2,4,10),labels=c("1/4","1/2","1","2","4","10"))+
            geom_text(aes(label=obs),vjust=-1,color="white")+
            geom_text(aes(label=paste("(",exp,")",sep="")),vjust=1,color="white")+
            labs(x="Chr 1",y="Chr 2",fill="O(Interactions)/E(Interactions)")

ggsave(chr_plot,file="data/dipterans/plots/D_melanogaster_chr_enrichment.pdf",width=12,height=7,dpi=400)
```

## Expression analysis

### Tissue

```{julia}


```

### Timing

## STRING database

The STRING database contains links between genes based on a variety of metrics. Here, we compare the co-evolution scores versus the STRING score, to see at what cut-off we are reaching all likely interactions.

```{julia}

string_db = CSV.read("data/dipterans/String_dmel.detailed.v12.0.txt",DataFrame)
string_alias = CSV.read("data/dipterans/String_dmel_protein_info.txt",DataFrame)


idx = indexin(gene_table.gene,string_alias.alias)
idx = replace(idx,nothing => missing)
gene_table.string_id = missings(String31,length(gene_table.gene))
gene_table.string_id[findall((!ismissing).(idx))] = string_alias.var"#string_protein_id"[idx[findall((!ismissing).(idx))]]

string_db.id1 = indexin(string_db.protein1,gene_table.string_id)
string_db.id2 = indexin(string_db.protein2,gene_table.string_id)
idx = intersect(findall((!isnothing).(string_db.id1)),findall((!isnothing).(string_db.id2)))
string_db = string_db[idx,:]
string_db.gid = [join(sort([gene_table.gene[x.id1],gene_table.gene[x.id2]]),"_") for x in eachrow(string_db)]


joint_tab = innerjoin(sig_ERC,string_db,on=:gid)
```

## Converting to a similarity matrix

For several types of analyses, we need to convert the long ERC output to a wide format in order to calculate similarity matrices.

```{julia}
using Tables
fERC_mat = missings(Float64,length(genes),length(genes))
for i in 1472:(length(genes)-1)
    idx_i_1 = findall(ERC.i .== gene_table.gene[i])
    idx_i_2 = findall(ERC.j .== gene_table.gene[i])
    for j in (i+1):length(genes)
        idx = union(findall(ERC.j[idx_i_1] .== gene_table.gene[j]),findall(ERC.i[idx_i_2] .== gene_table.gene[j]))
        if length(idx) == 1
            fERC_mat[i,j] = ERC.fERC[idx][1]
            fERC_mat[j,i] = ERC.fERC[idx][1]
        end
    end
end

CSV.write(data_dir*focal*"_fERC_matrix.tsv",Tables.table(fERC_mat),writeheader=false)

genes2 = union(sig_ERC.i,sig_ERC.j)
sig_mat = zeros(Float64,length(genes2),length(genes2))

for i in 1:(length(genes2)-1)
    idx_i_1 = findall(sig_ERC.i .== genes2[i])
    idx_i_2 = findall(sig_ERC.j .== genes2[i])
    for j in (i+1):length(genes2)
        idx = union(findall(sig_ERC.j[idx_i_1] .== genes2[j]),findall(sig_ERC.i[idx_i_2] .== genes2[j]))
        if length(idx) == 1
            sig_mat[i,j] = sig_ERC.fERC[idx][1]
            sig_mat[j,i] = sig_ERC.fERC[idx][1]
        end
    end
end

CSV.write(data_dir*focal*"_sigERC_matrix.tsv",Tables.table(sig_mat),writeheader=false)

```